Topological Equivariant Artist Model for Visualization Library Architecture

### 2022-08-17 paper/main.tex
why fb in human readable terms

% *** MATH PACKAGES ***
%% with other math-related packages, you may want to disable it.
\usepackage{amsmath, amsthm, amsfonts,amssymb,eulervm,xspace, mathtools}
\usepackage{stmaryrd}%mapsfrom 
\renewcommand{\restriction}{\mathord{\upharpoonright}} %restriction w/p space
\usepackage{mathrsfs} % math script fonts
\usepackage{relsize} %bigger
\usepackage{bm}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\theoremstyle{remark}
\newtheorem{example}{Example}[section]
% *** SPECIALIZED LIST PACKAGES ***

@@

\IEEEPARstart{V}isualizations are expected to be a faithful translation of data, meaning that inherent structure in the data should be present in the visualization; therefore the components of visualization software libraries that translate data to graphics are also expected to be structure preserving. This inherent structure can be quite complex, such as when the data is high dimensional, multivariate, deeply nested, or encoded in a distributed manner. We propose that we can rigorously specify what structure the library components are expected to preserve using category theory. We also propose that these specifications can generalize to most structure types by extending previous work on encoding  inherent structure in a consistent manner using mathematical structures from algebraic topology. These algebraic structures allow us to encode inherent structure as programmatic types, and the use of category theory facilitates developing a functional design framework that describes function composition in terms of functional algebraic operations. The typing system and composition inherent in a functional design encourage library developers to build complex visualization components from simple verifiable parts that have few side effects by design\cite{huHowFunctionalProgramming2015, hughesWhyFunctionalProgramming1989}. These categorically specified components could be built as a standalone library, but can also be integrated into existing libraries because they are inherently self-contained. Furthermore we see the application of these ideas in low level visualization library design as our motivation, and we hope these ideas will help shape the reorganization of critical data visualization libraries, such as Matplotlib. The contribution of this paper is a mathematical framework for describing structure preserving visualization components in a manner that is generalizable to different types of inherent structure. This framework provides also provides guidance for the construction and testing of structure preserving visualization library components.

@@

Visual algorithms assume the topology of their input data, as described in taxonomies of visualization algorithms Chi\cite{chiTaxonomyVisualizationTechniques2000} and by Troy and M\"{o}ller \cite{toryRethinkingVisualizationHighlevel2004}. For example, a \texttt{line} algorithm often does not have a way to query whether a list of (x,y) coordinates is the distinct rows, the time series, or the list of stations in \autoref{fig:related-work:continuity:ktypes}. While plotting the time series as a continuous line would be correct, it would be incorrect for a visualization to indicate that the distinct rows or stations are connected in a 1D continuous manner. Since topological continuity is generalizable, we propose the construction of topology types such that data can explicitly state its topology type and visual algorithms can check that this type matches their assumptions. The typing system proposed in this paper builds on Butler's proposal of using a mathematical structure called fiber bundles as an abstract data representation in visualization \cite{butlerVectorBundleClassesForm1992, butlerVisualizationModelBased1989}. We extend Butler's work because he discusses how bundles can be used to encode data in a consistent manner for a number of different \note{arbitrary?} topological and field structures. 
 We sketch out fiber bundles in \autoref{sec:atct:fiber-bundles}, but his work provides a thorough introduction to bundles for visualization practitioners.

\subsection{Data Field Structure}
\label{sec:related-work:equivariance}
Along with preserving topology, visualization algorithms are expected to preserve the structure of the data fields, i.e. the columns in a table, when they translate data to visual representations. 
The field structure of data has traditionally been described using the measurment scales-i.e. 
nominal, ordinal, interval, ratio - which Steven's classified by their mathematical group structure\cite{stevensTheoryScalesMeasurement1946a} and others have formally expanded to include more types \cite{leaFormalizationMeasurementScale, thomasMathematizationNotMeasurement2014}. The expectation that visual transformation is one where the measurement structure be matched by the visual encoding was codified by Bertin \cite{bertinSemiologyGraphicsDiagrams2011}. Two mathematical concepts for expressing structure preservation are \textit{equivariance} and \textit{homomorphism}:
\begin{itemize}
\item an \textit{equivariant} function preserves symmetric group structure on its input and output. Given a group $G$ that acts on input $X$ and output $Y$ and a function $f: X \rightarrow Y$, the function $f$ is \textit{equivariant} if $f(g(x)) = g(f(x))$ for all $g$ in $G$ and for all $x$ in $X$\cite{nlab:equivariant}.
\item a \textit{homomorphic} function preserves binary operations on input and output of the same algberaic type. A map $f: X \rightarrow Y$ is a \textit{homomorphism} when $X$ and $Y$ have the same algebraic structure type and $f(x_1 \oslash x_2) = f(x_1) \oslash f(x_2)$ for every pair $x_1, x_2$ in $X$\cite{grimaldiDiscreteCombinatorialMathematics2006}.
\end{itemize}
For example, nominal measurements are permutable, meaning that they are acted on by elements of a permutation group. For the visual transformation to be equivariant, that nominal measurement must be mapped to a visual encoding that is also permutable, such as distinct marker shapes or color hues. Partial orders are non-invertable, so a visual encoding is homomorphic when data that is partially ordered is mapped to visual encodings that are partially ordered in the same way.Mackinlay\cite{mackinlayAutomaticDesignGraphical1987} proposes that data to visual encoders should be \textit{homomorphic} maps in his specification of \textit{A Presentation Tool}. The notion that visual transforms are equivariant underlies Kindlemannn and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014}'s Algebraic Visualization Design. They propose three specific types of structure preservation: that the visualization should not change if the data representation (i.e. the data container) changes, that data maps unambiguously to visual elements \cite{ziemkiewiczEmbeddingInformationVisualization2009}, and that changing the data should correspond to changes in the visualization in a perceptually significant manner. In \autoref{sec:atct:sheaves} we introduce a uniform abstract data representation layer, the expectation of unambiguous mappings is expressed in \autoref{sec:atct:xi}, and the definition of equivariance introduced in \autoref{sec:artist:equivariant:artist} includes the correspondence between data and visual changes. In this work, we formally specify the conditions necessary to build structure preserving components using category theory, because it provides a rich language for expressing the structure of objects, functions between objects, and the constraints that must hold for these functions to compose \cite{wielsManagementEvolvingSpecifications1998,yorgeyMonoidsThemeVariations}. A brief visualization oriented introduction to category theory is in Vickers et al \cite{vickersUnderstandingVisualizationFormal2013}, but they are applying category theory to semantic concerns about visualization design rather than library architecture.

\subsection{Structure Preservation In Software}
\label{sec:related-work:software}
Visualization libraries are in part measured by how expressive the components of the library are, where expressiveness is a measure of which structure preserving mappings a tool can implement \cite{mackinlayAutomatingDesignGraphical1986}. While some visualization tools aim to automate the pairing of data with structure preserving visual representations, such as Tableau\cite{StoltePolaris2002,hanrahanVizQL2006,MackinlayShowme2007}, many visualization libraries leave that choice to the user. For example, connectivity assumptions tend to be embedded in each of the visual algorithms of `building block` libraries, a term coined by by Wongsuphasawat \cite{wongsuphasawatNavigatingWideWorld2021} to describe libraries that provide modular components for building elements of a visualization, such as functions for making boxes or translating data values to colors. In building block libraries such as Matplotlib\cite{hunterMatplotlib2DGraphics2007} and D3\cite{bostockDataDrivenDocuments2011} the connectivity assumptions lead to very different interfaces; for example in Matplotlib methods for updating the data in a plot differ across the various plotting objects and the parameter list is inconsistent as to which parameters can be applied per data point or for the whole element, and in D3 the API for network visualizers is mostly distinct from the rest of the modules. While VTK\cite{hanwellVisualizationToolkitVTK2015,geveciVTK2012} provides a language for expressing the topological properties of the data, and therefore can embed that information in its visual algorithms, it does so in a non-uniform manner. On the other hand, domain specific libraries are designed with the assumption of continuities that are common in the domain \cite{HeerSoftware2006}, and therefore can somewhat restrict the interface to choices that that are appropriate for the domain. For example, a tabluar topological structure of discret rows, as illustrated in \autoref{fig:related-work:continuity:ktypes}, is assumed by 
A Presentation Tool\cite{mackinlayAutomatingDesignGraphical1986, mackinlayAutomatingDesignGraphical1986} and grammar of graphics\cite{wilkinsonGrammarGraphics2005} and the ggplot\cite{wickhamGgplot2ElegantGraphics2016}, vega\cite{satyanarayanDeclarativeInteractionDesign2014}, and altair\cite{vanderplasAltairInteractiveStatistical2018} libraries built on these frameworks. Image libraries such as Napari\cite{nicholas_sofroniew_2021_4533308} and ImageJ\cite{schneiderNIHImageImageJ2012} and its humanities ImagePlot\cite{studiesCulturevisImageplot2021} plugin assume that the input is 2D continuous. Networking libraries such as gephi\cite{bastianGephiOpenSource2009} and networkx\cite{HagbergExploringNetwork2008} assume a graph-like structure. By assuming the structure of their data, these domain specific libraries can provide more cohesive interfaces for a much more limited set of visualization algorithms than the building block libraries offer.

@@

\begin{definition} A \textbf{fiber bundle} $(\dtotalc, \dbasec, \pi, \dfiberc)$ is a structure with topological spaces $\dtotalc, \dfiberc, \dbasec$ and continuous surjective map $\pi: \dtotalc \rightarrow \dbasec$ \cite{FiberBundle2020,spanier1989algebraic}. 
\begin{equation}
  \label{eq:atct:fb:intro}
  \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
   \dfiberc
    \arrow[r, hook, color=total] \& 
    \dtotalc
    \arrow[d, "\pi"',color=total] \\
     \& 
  \dbasec
     \arrow[u, "\dsectionc"', bend right, pos=.5, color=section, dashed]
  \end{tikzcd}
\end{equation} 
A map $\pi$ is a bundle projection map when 
\begin{enumerate}
  \item the \textcolor{fiber}{fiber space} \dfiberc\ is the preimage of the projection function $\pi$ at a point $\dbasepoint$ in the \textcolor{base}{base space} \dbasec\ such that $\dfiber_{\dbasepoint} = \pi^{-1}(\dbasepoint)$
  \item there is an open neighborhood $\openset_{\dbasepoint}$ surrounding each point in the base space $\dbasepoint \in \openset_{\dbasepoint} \subset \dbase$ such that the \textcolor{total}{total space} \dtotalc\ over the neighborhood is locally trivial $\dtotal\restriction_{\openset} = \openset \times \dfiber$
\end{enumerate}
\end{definition}
The first condition for a fiber bundle means that all points in a fiber space over a point in the base space map to the same base space point, which we rely on to establish that there is a formal binding between typed data value, discussed in \autoref{sec:atct:fb:fiber} and the corresponding index, discussed in \autoref{sec:atct:fb:base}. The condition of local triviality means that the fibers are identical over open neighborhoods and is what allows fiber bundles to be extremely generalizable, as discussed in \autoref{sec:atct:fb:sections}

\begin{definition} A \textcolor{section}{\textbf{section}} $\dsectionc: \dbasec \rightarrow \dtotalc$ over a fiber bundle is a smooth right inverse of $\pi$ such that $\pi(\dsection(\dbasepoint)) = \dbasepoint$ for all $\dbasepoint \in \dbase$
\end{definition}
In \autoref{sec:atct:fb:sections}, we propose that modeling data as sections provides a way to encapsulate topological and field structure in a uniform dimension and type independent manner.

@@

The topological structure of the data can be expressed programmatically by constructing data types that encapsulate the class of topological structure-.i.e point, line, plane, network, etc. We propose that these data types can be formally specified as objects of a category. 
\begin{definition} 
   An arbitrary \textbf{category} $\mathcal{C}$ consists of the following \textit{data}: 
\begin{enumerate}
  \item \textit{object}: X
  \item \textit{identity morphism}: $X \xrightarrow {id_x} X$
  \item \textit{morphisms}: $X \xrightarrow{f} Y$ between objects X, Y in $\mathcal{C}$
  \item \textit{composite morphism}: given the morphisms $X \xrightarrow{f} Y$ and $Y \xrightarrow{g} Z$, there exists a morphism $X \xrightarrow{g\circ f} Z$
\end{enumerate}
The data must satisfy the following \textit{rules}
\begin{enumerate}
  \item \textit{unitality:} for every morphism $ X \xrightarrow{f} Y$, there exists identity morphisms $id_x, id_y$ such that $f \circ id_x = f = id_y \circ f$
  \item \textit{associativity:} if any three morphisms $f, g, h$ are composable, 
    \begin{equation*}
      \begin{tikzcd}
        X \arrow[r, "f"] \arrow[rrr, "h\circ(g\circ f) = (h\circ g)\circ f"', bend right, dashed] & Y  \arrow[r, "g"] & Z \arrow[r, "h"] & W
        \end{tikzcd}
  \end{equation*}
  then they are associative such that $h\circ(g\circ f) = (h \circ g) \circ f$  \cite{lawvere2009conceptual,riehlCategoryTheoryContext,maclaneCategoriesWorkingMathematician2013,fongInvitationAppliedCategory2019}. 
  \end{enumerate}
\end{definition}

@@

\subsubsection{Uniform Abstract Graphic Representation}
One of the advantages of fiber bundles is that they are general enough that we can also encode the output of a visual algorithm as a bundle. This allows us to use the same structure to express the properties of data and the graphic that must be symmetric to the data in an equivariant (\autoref{sec:related-work:equivariance}) transformation. We denote the output as a graphic, but the use of bundles allows us to generalize to output on any display space, such as a screen or 3D print. 
\begin{equation}
  \label{eq:atct:fb:graphic}
  \gfiberc \hookrightarrow \gtotalc \xrightarrow{\pi} \gbasec
\end{equation}
The total space \gtotalc\ is an abstraction of an ideal (infinite resolution) space into which the graphic can be rendered. The base space \gbasec\ is a parameterization of the display area, for example the inked bounding box in cairo \cite{CairographicsOrg}. The fiber space \gfiberc\ is an abstraction of the renderer fields, for example $\{x,\,y,\,r,\,g,\,b,\,a\}$.

>>>>

% *** MATH PACKAGES ***
%% with other math-related packages, you may want to disable it.
\usepackage{amsmath, amsthm, amsfonts,amssymb,eulervm,xspace, mathtools}
\usepackage{stmaryrd}%mapsfrom 
%\renewcommand{\restriction}{\mathord{\upharpoonright}} %restriction w/p space
\usepackage{mathrsfs} % math script fonts
\usepackage{relsize} %bigger
\usepackage{bm}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\theoremstyle{remark}
\newtheorem{example}{Example}[section]
% *** SPECIALIZED LIST PACKAGES ***

@@

\IEEEPARstart{V}isualizations are expected to be a faithful translation of data, meaning that inherent structure in the data should be present in the visualization; therefore the components of visualization software libraries that translate data to graphics are also expected to be structure preserving. This inherent structure can be quite complex, such as when the data is high dimensional, multivariate, deeply nested, or encoded in a distributed manner. We propose that we can rigorously specify what structure the library components are expected to preserve using category theory. We also propose that these specifications can generalize to most structure types by extending previous work on encoding  inherent structure in a consistent manner using mathematical structures from algebraic topology. These algebraic structures can be translated into analogous programmatic types for encoding structure, while the use of category theory facilitates developing a functional design framework that describes function composition in terms of functional algebraic operations. The typing system and composition inherent in a functional design encourage library developers to build complex visualization components from simple verifiable parts that have few side effects \cite{huHowFunctionalProgramming2015, hughesWhyFunctionalProgramming1989}. These categorically specified components could be built as a standalone library and integrated into existing libraries because they are inherently self-contained. Furthermore we see the application of these ideas in low level visualization library design as our motivation, and we hope these ideas will help shape the reorganization of critical data visualization libraries, such as Matplotlib. The contribution of this paper is a mathematical framework for describing structure preserving visualization components in a manner that is generalizable to different types of inherent structure. This framework also provides guidance for the construction and testing of structure preserving visualization library components.

@@

Visual algorithms assume the topology of their input data, as described in taxonomies of visualization algorithms Chi\cite{chiTaxonomyVisualizationTechniques2000} and by Troy and M\"{o}ller \cite{toryRethinkingVisualizationHighlevel2004}. For example, a \texttt{line} algorithm often does not have a way to query whether a list of (x,y) coordinates is the distinct rows, the time series, or the list of stations in \autoref{fig:related-work:continuity:ktypes}. While plotting the time series as a continuous line would be correct, it would be incorrect for a visualization to indicate that the distinct rows or stations are connected in a 1D continuous manner. Since topological continuity is generalizable, we propose the construction of topology types such that data can explicitly state its topology type and visual algorithms can check that this type matches their assumptions. The typing system proposed in this paper builds on Butler's proposal of using a mathematical structure called fiber bundles as an abstract data representation in visualization \cite{butlerVectorBundleClassesForm1992, butlerVisualizationModelBased1989}. We extend Butler's work because he discusses how bundles can be used to encode data in a consistent manner for the topological and field structures that are common in visualization. We sketch out fiber bundles in \autoref{sec:atct:fiber-bundles}, but his work provides a thorough introduction to bundles for visualization practitioners.

\subsection{Field Structure} 
\label{sec:related-work:equivariance}
As with topology, Bertin \cite{bertinSemiologyGraphicsDiagrams2011} codified the expectation that a data value and its visual encoding are expected to have compatible structure. The structure on values has traditionally been described  using the measurement scales-i.e. nominal, ordinal, interval, ratio - which Steven's classified by their mathematical group structure\cite{stevensTheoryScalesMeasurement1946a} and others have formally expanded to include more types \cite{leaFormalizationMeasurementScale, thomasMathematizationNotMeasurement2014}. Generally the structure on each field of the dataset is preserved separately, i.e. each column in a table is mapped to a distinct encoding. Two mathematical concepts for describing a function that preserves structure are \textit{equivariance} and \textit{homomorphism}. An \textit{equivariant} function preserves symmetric group structure on its input and output. Given a group $G$ that acts on input $X$ and output $Y$ and a function $f: X \rightarrow Y$, the function $f$ is \textit{equivariant} when $f(g(x)) = g(f(x))$ for all $g$ in $G$ and for all $x$ in $X$\cite{nlab:equivariant}. A \textit{homomorphic} function $f$ preserves binary operations on input and output of the same algebraic type. A map $f: X \rightarrow Y$ is a \textit{homomorphism} when  $f(x_1 \oslash x_2) = f(x_1) \oslash f(x_2)$ for every pair $x_1, x_2$ in $X$\cite{grimaldiDiscreteCombinatorialMathematics2006}. For example, nominal measurements are permutable, meaning that they are acted on by elements of a permutation group. For the visual transformation to be equivariant, that nominal measurement must be mapped to a visual encoding that is also permutable, such as distinct marker shapes or color hues. Partial orders are non-invertable, so a visual encoding is homomorphic when data that is partially ordered is mapped to visual encodings that are partially ordered in the same way.

The notion that visual encodings should be \textit{homomorphic} was proposed by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987} his specification of \textit{A Presentation Tool} and the idea that visual transforms are \textit{equivariant} underlies Kindlemann and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014}'s Algebraic Visualization Design. Kindlemann and Scheidegger propose three specific types of structure preservation: that the visualization should not change if the data representation (i.e. the data container) changes, that data maps unambiguously to visual elements \cite{ziemkiewiczEmbeddingInformationVisualization2009}, and that changing the data should correspond to changes in the visualization in a perceptually significant manner. In \autoref{sec:atct:sheaves} we introduce a uniform abstract data representation layer, the expectation of unambiguous mappings is expressed in \autoref{sec:atct:xi}, and the definition of equivariance introduced in \autoref{sec:artist:equivariant:artist} includes the correspondence between data and visual changes. In this work, we formally specify the conditions necessary to build structure preserving components using category theory, because it provides a rich language for expressing the structure of objects, functions between objects, and the constraints that must hold for these functions to compose \cite{wielsManagementEvolvingSpecifications1998,yorgeyMonoidsThemeVariations}. A brief visualization oriented introduction to category theory is in Vickers et al \cite{vickersUnderstandingVisualizationFormal2013}, but they are applying category theory to semantic concerns about visualization design rather than library architecture.

\subsection{Structure Preservation In Software}
\label{sec:related-work:software}
Visualization libraries are in part measured by how expressive the components of the library are, where expressiveness is a measure of which structure preserving mappings a tool can implement \cite{mackinlayAutomatingDesignGraphical1986}. While some visualization tools aim to automate the pairing of data with structure preserving visual representations, such as Tableau\cite{StoltePolaris2002,hanrahanVizQL2006,MackinlayShowme2007}, many visualization libraries leave that choice to the user. For example, connectivity assumptions tend to be embedded in each of the visual algorithms of `building block` libraries, a term used by Wongsuphasawat \cite{wongsuphasawatNavigatingWideWorld2021} to describe libraries that provide modular components for building elements of a visualization, such as functions for making boxes or translating data values to colors. In building block libraries such as Matplotlib\cite{hunterMatplotlib2DGraphics2007} and D3\cite{bostockDataDrivenDocuments2011} the connectivity assumptions lead to very different interfaces; for example in Matplotlib methods for updating data and parameters for controlling aesthetics differ between plots. While VTK\cite{hanwellVisualizationToolkitVTK2015,geveciVTK2012} provides a language for expressing the topological properties of the data, and therefore can embed that information in its visual algorithms, it does so in a non-uniform manner. On the other hand, domain specific libraries are designed with the assumption of continuities that are common in the domain \cite{HeerSoftware2006}, and therefore can somewhat restrict the interface to choices that that are appropriate for the domain. For example, a tabular topological structure of discret rows, as illustrated in \autoref{fig:related-work:continuity:ktypes}, is assumed by 
A Presentation Tool\cite{mackinlayAutomatingDesignGraphical1986, mackinlayAutomatingDesignGraphical1986} and grammar of graphics\cite{wilkinsonGrammarGraphics2005} and the ggplot\cite{wickhamGgplot2ElegantGraphics2016}, vega\cite{satyanarayanDeclarativeInteractionDesign2014}, and altair\cite{vanderplasAltairInteractiveStatistical2018} libraries built on these frameworks. Image libraries such as Napari\cite{nicholas_sofroniew_2021_4533308} and ImageJ\cite{schneiderNIHImageImageJ2012} and its humanities ImagePlot\cite{studiesCulturevisImageplot2021} plugin assume that the input is 2D continuous. Networking libraries such as gephi\cite{bastianGephiOpenSource2009} and networkx\cite{HagbergExploringNetwork2008} assume a graph-like structure. By assuming the structure of their data, these domain specific libraries can provide more cohesive interfaces for a much more limited set of visualization algorithms than the building block libraries offer.

@@

\begin{definition} A \textbf{fiber bundle} $(\dtotalc, \dbasec, \pi, \dfiberc)$ is a structure with topological spaces $\dtotalc, \dfiberc, \dbasec$ and  bundle projection map $\pi: \dtotalc \rightarrow \dbasec$ \cite{FiberBundle2020,spanier1989algebraic}. 
\begin{equation}
  \label{eq:atct:fb:intro}
  \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
   \dfiberc
    \arrow[r, hook, color=total] \& 
    \dtotalc
    \arrow[d, "\pi"',color=total, two heads] \\
     \& 
  \dbasec
     \arrow[u, "\dsectionc"', bend right, pos=.5, color=section, dashed]
  \end{tikzcd}
\end{equation} 
A continuous surjective map \textbf{$\pi$ is bundle projection} map when 
\begin{enumerate}
  \item the \textcolor{fiber}{fiber space} \dfiberc\ is the preimage of the projection function $\pi$ at a point $\dbasepoint$ in the \textcolor{base}{base space} \dbasec\ such that $\dfiber_{\dbasepoint} = \pi^{-1}(\dbasepoint)$. All fibers in a bundle are isomorphic such that $\dfiber \cong \dfiber_{\dbasepoint}$ for all points $\dbasepoint \in \dbase$. 
  \item there is an open neighborhood $\openset_{\dbasepoint}$ surrounding each point in the base space $\dbasepoint \in \openset_{\dbasepoint} \subset \dbase$ such that the \textcolor{total}{total space} \dtotalc\ over the neighborhood, denoted $\dtotal\restriction\openset$\footnote{The symbol $\restriction$ is the restriction operator\cite{RestrictionMathematics2022}. For example $\pi^{-1}\restriction\openset$ is the inverse function $\pi^{-1}$ only defined on $\openset$}, is locally trivial. The condition for local triviality is that $\dtotal\restriction_{\openset} = \openset \times \dfiber = \pi^{-1}\restriction_{\openset}$
\end{enumerate}
\end{definition}
We propose that the map $\pi$ can express the formal binding between having a typed data field, discussed in \autoref{sec:atct:fb:fiber} and a corresponding index, as discussed in \autoref{sec:atct:fb:base}. For our purposes, the advantage of fiber bundles is that the field type and topological structure are unconstrained in terms of dimensionality and the only conditions that must be satisfied are that every index point has a corresponding field type and the field types must be the same (isomorphic) for every index point.

\begin{definition} A \textcolor{section}{\textbf{section}} $\dsectionc: \dbasec \rightarrow \dtotalc$ over a fiber bundle is a smooth right inverse of $\pi$ such that $\pi(\dsection(\dbasepoint)) = \dbasepoint$ for all $\dbasepoint \in \dbase$
\end{definition}
We propose that modeling data as sections provides a way to encapsulate topological and field structure in a uniform dimension and type independent manner, as discussed in \autoref{sec:atct:fb:sections}.

@@

The topological structure of the data can be expressed programmatically by constructing data types that encapsulate the class of topological structure-.i.e point, line, plane, network, etc. We propose that these data types can be formally specified as objects of a category. 
\begin{definition} 
   An \textbf{category} $\mathcal{C}$ consists of the following \textit{data}: 
\begin{enumerate}
  \item \textit{object}: X
  \item \textit{identity morphism}: $X \xrightarrow {id_x} X$
  \item \textit{morphisms}: $X \xrightarrow{f} Y$ between objects X, Y in $\mathcal{C}$
  \item \textit{composite morphism}: given the morphisms $X \xrightarrow{f} Y$ and $Y \xrightarrow{g} Z$, there exists a morphism $X \xrightarrow{g\circ f} Z$
\end{enumerate}
The data must satisfy the following \textit{rules}
\begin{enumerate}
  \item \textit{unitality:} for every morphism $ X \xrightarrow{f} Y$, there exists identity morphisms $id_x, id_y$ such that $f \circ id_x = f = id_y \circ f$
  \item \textit{associativity:} if any three morphisms $f, g, h$ are composable, 
    \begin{equation*}
      \begin{tikzcd}
        X \arrow[r, "f"] \arrow[rrr, "h\circ(g\circ f) = (h\circ g)\circ f"', bend right, dashed] & Y  \arrow[r, "g"] & Z \arrow[r, "h"] & W
        \end{tikzcd}
  \end{equation*}
  then they are associative such that $h\circ(g\circ f) = (h \circ g) \circ f$  \cite{lawvere2009conceptual,riehlCategoryTheoryContext,maclaneCategoriesWorkingMathematician2013,fongInvitationAppliedCategory2019}. 
  \end{enumerate}
\end{definition}

@@

\subsubsection{Uniform Abstract Graphic Representation}
One of the advantages of fiber bundles is that they are general enough that we can also encode the output of a visual algorithm as a bundle. This allows us to use the same structure to express the properties of data and the graphic that must be symmetric to the data in an equivariant (\autoref{sec:related-work:equivariance}) transformation. We denote the output as a graphic, but the use of bundles allows us to generalize to output on any display space, such as a screen or 3D print. 
\begin{equation}
  \label{eq:atct:fb:graphic}
  \begin{tikzcd}
      \gfiberc \arrow[r, hook, color=total] & \gtotalc \arrow[r, "\pi", two heads, color=total] & \gbasec
  \end{tikzcd}
\end{equation}
The total space \gtotalc\ is an abstraction of an ideal (infinite resolution) space into which the graphic can be rendered. The base space \gbasec\ is a parameterization of the display area, for example the inked bounding box in cairo \cite{CairographicsOrg}. The fiber space \gfiberc\ is an abstraction of the renderer fields, for example $\{x,\,y,\,r,\,g,\,b,\,a\}$.
